Multi-level modeling with Openflexo/\FML

### 2021-10-07 8-discussion.tex
synchro
des questions + maj

  The Openflexo approach is level-agnostic: ``levels'' have no specific nature
  and there are no numbered levels. In our solution, concepts of a given level
  are grouped into a virtual model. Inheritance and instantiation allow the
  establishment of relationships between concepts from different levels.
%\ref{fig:BPMNSubsetExample}
%  Figure~\ref{fig:AcmeFullArchitecture} shows three levels: the metamodel, the
%  ACME metamodel and the ACME model.
Figure~\ref{fig:MultilevelArchitecture} shows the multilevel architecture of
our solution.
\noteJC{updated}

@@

  Due to the fact that our approach is level-agnostic, our solution could have
  more or fewer levels depending on the variations of the use case. However,
  the number of levels is related to the problem, making the approach fixed (to
  3).

@@

  Thanks to the level-agnosticism nature of the Openflexo approach, cross-level
  relationships are not an issue. Model elements of different levels can be
  linked each other in a transparent way, using inheritance or instantiation.
  For instance in the figure~\ref{fig:AcmeFullArchitecture}, the
  \texttt{Developer} concept from \texttt{Acme} metamodel specializes the
  \texttt{ActorType} of the base metamodel, and the \texttt{Developer} of the
  \texttt{Acme} model is an instance of the metamodel concept.
%  \textcolor{red}{Phrase faite a relire sur Developer\dots}
\noteJC{updated}

@@

  Behaviors are continuously checked, ensuring the integrity of the models when
  changes occur. However, most of these behaviors have to be written by the
  users. Thus, the integrity of the contents relies on the users when they
  build the metamodels and the models. The section~\ref{sec:model} describes
  how the requirements are captured and how behaviors are implemented with
  \FML: for instance, \texttt{isAuthorizedActor} behavior is needed to enforce
  \textbf{P17} requirement, leading to writing \FML code in \texttt{TaskType}
  and \texttt{Task}. Note that in our approach, the metamodels are built in an
  ad hoc way together with the models (co-construction), therefore the user
  also validates on an ad hoc basis.
\noteJC{updated}

@@

  Due to the nature of our approach and its level-agnosticism, \emph{deep
  characterization} does not apply to our solution. Such a mechanism could
  probably be encoded by specific concepts and constraints (behaviors).
  However, it would not be a generic mechanism, making it difficult to reuse
  for another problem.  Plus, programming the notion of \emph{depth} would
  probably imply that a virtual model knows its sub-elements, which is not
  desired and considered as bad design.
\noteJC{updated}

>>>>

  The Openflexo approach is level-agnostic: ``levels'' have no specific nature
  and there are no numbered levels. In our solution, concepts of a given level
  are grouped into a virtual model. Inheritance and instantiation allow the
  establishment of relationships between concepts from different levels.
%\ref{fig:BPMNSubsetExample}
%  Figure~\ref{fig:AcmeFullArchitecture} shows three levels: the metamodel, the
%  ACME metamodel and the ACME model.
Figure~\ref{fig:MultilevelArchitecture} shows the multilevel architecture of
our solution.

@@

  Due to the fact that our approach is level-agnostic, our solution could have
  more or fewer levels depending on the variations of the use case. However,
  the number of levels is related to the problem, our solution uses 3 levels.

@@

  Thanks to the level-agnosticism nature of the Openflexo approach, cross-level
  relationships are not an issue. Model elements of different levels can be
  linked each other in a transparent way, using inheritance or instantiation.
  For instance in the figure~\ref{fig:AcmeFullArchitecture}, the
  \texttt{Developer} concept from \texttt{Acme} metamodel specializes the
  \texttt{ActorType} of the base metamodel, and the \texttt{Developer} {\color{red}?} of the
  \texttt{Acme} model is an instance of the metamodel concept.
%  \textcolor{red}{Phrase faite a relire sur Developer\dots}

@@

  Behaviors are continuously checked, ensuring the integrity of the models when
  changes occur. However, most of these behaviors have to be written by the
  users. Thus, the integrity of the contents relies on the users when they
  build the metamodels and the models. The section~\ref{sec:model} describes
  how the requirements are captured and how behaviors are implemented with
  \FML. For instance, \texttt{isAuthorizedActor} behavior is needed to enforce
  \textbf{P17} requirement, leading to writing \FML code in \texttt{TaskType}
  and \texttt{Task}. Note that in our approach, the metamodels are built in an
  {\color{red}ad hoc} way together with the models (co-construction), therefore the user
  also validates on an {\color{red}ad hoc} basis.

@@

  Due to the nature of our approach and its level-agnosticism, \emph{deep
  characterization} does not apply to our solution. Such a mechanism could
  probably be encoded by specific concepts and constraints (behaviors).
  However, it would not be a generic mechanism, making it difficult to reuse
  for another problem.  Plus, programming the notion of \emph{depth} would
  probably imply that a virtual model knows its sub-elements, which is not
  desired and considered as bad design.
